Gravitational wave sources in our Galactic backyard: Predictions for BH and NS binaries in LISA

### 2021-07-21 paper/tex/sections/conclusion.tex
update numbers

Our main conclusions can be summarised as:
\begin{itemize}
    \item \textbf{Detection rate:} We find that on average, for our fiducial model, a 4(10)-year LISA mission will detect \BHBHFourYear{}(\BHBHTenYear{}) BHBHs, \BHNSFourYear{}(\BHNSTenYear{}) BHNSs and \NSNSFourYear{}(\NSNSTenYear{}) NSNSs.
    \item \textbf{Black hole mass distribution:} We find that the black hole mass distribution for both BHBHs and BHNSs that are detectable with LISA are not skewed towards heavier masses (as in ground-based detectors), with 88\% and 91\% having masses less than $11 \unit{M_\odot}$ respectively. Moreover, we find that for LISA detectable DCOs, 35\% of BHBHs and 39\% of BHNSs have masses in the theoretical lower mass gap between NSs and BHs. This implies that LISA will be able provide conclusive evidence about the existence of a lower-mass gap. 
    \item \textbf{Eccentricity distribution:} We find that BH and NS binaries still have significant eccentricities when observable by the LISA mission, unlike the more numerous WDWD LISA population. We find that 86(20)\% of BHBHs, 46(8)\% of BHNSs and 86(12)\% of NSNSs have eccentricity of $e > 0.01(0.3)$ at the start of the LISA mission.
    \item \textbf{Physics variations:} For BHBHs, we find that the detection rate is largely unaffected by changes in underlying physics assumptions, except for the optimistic CE scenario and increased Wolf-Rayet winds. Conversely, the BHNS rate varies widely with physics variation, ranging from a mean of 4 to 138. The NSNS shows some variation but only shows large changes under the assumption that case BB mass transfer is unstable and that core-collapse supernovae produce weaker kicks.
    \item \textbf{Distinguishable sources:} For a 4(10)-year LISA mission, we estimate that of the detectable population, at least \BHBHNotWDWDFour{}(\BHBHNotWDWDTen{}) BHBHs, \BHNSNotWDWDFour{}(\BHNSNotWDWDTen{}) BHNSs and \NSNSNotWDWDFour{}(\NSNSNotWDWDTen{}) NSNSs will produce signals that are distinguishable from a signal produced by a WDWD. Additionally, we predict that we will be able to determine whether \BHBHDistinguishedFour{}(\BHBHDistinguishedTen{}) BHBHs, \BHNSDistinguishedFour{}(\BHNSDistinguishedTen{}) BHNSs and \NSNSDistinguishedFour{}(\NSNSDistinguishedTen{}) NSNSs are signals from a binary that contains a black hole.
    \item \textbf{Joint SKA-LISA detections:} We expect that \textit{if} the binary contains a pulsar that is beaming towards Earth, SKA-1 will be able to detect at least 10 LISA DCOs that contain a NS, whilst the increased number of pulsars that could crowd the sky in SKA-2 means that this total decreases to 6. This total could be much higher if we consider that SKA and LISA could match their orbital frequency estimates to select the correct pulsar from a crowded field.
    \item \textbf{Importance of choice of MW model:} Many studies use Milky Way models that use fixed metallicity populations which are assigned irrespective of birth time or position, do not account for the inside-out growth of the thin disc and use constant star formation rates. The use of these simpler MW models may lead to an overestimation of the LISA NSNS detection rate. It may also introduce unphysical artifacts into DCO parameter distributions, particularly the mass distributions, which lead to inaccurate predictions.
\end{itemize}

>>>>

Our main conclusions can be summarised as:
\begin{itemize}
    \item \textbf{Detection rate:} We find that on average, for our fiducial model, a 4(10)-year LISA mission will detect \BHBHFourYear{}(\BHBHTenYear{}) BHBHs, \BHNSFourYear{}(\BHNSTenYear{}) BHNSs and \NSNSFourYear{}(\NSNSTenYear{}) NSNSs.
    \item \textbf{Black hole mass distribution:} We find that the black hole mass distribution for both BHBHs and BHNSs that are detectable with LISA are not skewed towards heavier masses (as in ground-based detectors), with $\BHBHmBHBelowEleven{}$ and $\BHNSmBHBelowEleven{}$ having masses less than $11 \unit{M_\odot}$ respectively. Moreover, we find that for LISA detectable DCOs, $\BHBHLowerMassGap{}$ of BHBHs and $\BHNSLowerMassGap{}$ of BHNSs have masses in the theoretical lower mass gap between NSs and BHs. This implies that LISA will be able provide conclusive evidence about the existence of a lower-mass gap. 
    \item \textbf{Eccentricity distribution:} We find that BH and NS binaries still have significant eccentricities when observable by the LISA mission, unlike the more numerous WDWD LISA population. We find that \BHBHNotCirc{}(\BHBHHighlyEccentric{}) of BHBHs, \BHNSNotCirc{}(\BHNSHighlyEccentric{}) of BHNSs and \NSNSNotCirc{}(\NSNSHighlyEccentric{})\% of NSNSs have eccentricity of $e > 0.01(0.3)$ at the start of the LISA mission.
    \item \textbf{Physics variations:} For BHBHs, we find that the detection rate is largely unaffected by changes in underlying physics assumptions, except for the optimistic CE scenario and increased Wolf-Rayet winds. Conversely, the BHNS rate varies widely with physics variation, ranging across three orders of magnitude. The NSNS shows some variation but only shows large changes under the assumption that case BB mass transfer is unstable and that core-collapse supernovae produce weaker kicks.
    \item \textbf{Distinguishable sources:} For a 4(10)-year LISA mission, we estimate that of the detectable population, at least \BHBHNotWDWDFour{}(\BHBHNotWDWDTen{}) BHBHs, \BHNSNotWDWDFour{}(\BHNSNotWDWDTen{}) BHNSs and \NSNSNotWDWDFour{}(\NSNSNotWDWDTen{}) NSNSs will produce signals that are distinguishable from a signal produced by a WDWD. Additionally, we predict that we will be able to determine whether \BHBHDistinguishedFour{}(\BHBHDistinguishedTen{}) BHBHs, \BHNSDistinguishedFour{}(\BHNSDistinguishedTen{}) BHNSs and \NSNSDistinguishedFour{}(\NSNSDistinguishedTen{}) NSNSs are signals from a binary that contains a black hole.
    \item \textbf{Joint SKA-LISA detections:} We expect that \textit{if} the binary contains a pulsar that is beaming towards Earth, SKA-1 will be able to detect at least 10 LISA DCOs that contain a NS, whilst the increased number of pulsars that could crowd the sky in SKA-2 means that this total decreases to 6. This total could be much higher if we consider that SKA and LISA could match their orbital frequency estimates to select the correct pulsar from a crowded field.
    \item \textbf{Importance of choice of MW model:} Many studies use Milky Way models that use fixed metallicity populations which are assigned irrespective of birth time or position, do not account for the inside-out growth of the thin disc and use constant star formation rates. The use of these simpler MW models may lead to an overestimation of the LISA NSNS detection rate. It may also introduce unphysical artifacts into DCO parameter distributions, particularly the mass distributions, which lead to inaccurate predictions.
\end{itemize}

### 2021-07-21 paper/tex/sections/results.tex
update numbers

\subsubsection{Black Hole Mass}
For both the BHBHs and BHNSs, the black hole mass distribution extends to relatively low masses, with $88\%$ and $90\%$ respectively below $11 \unit{M_{\odot}}$. Unlike ground-based detectors, LISA is not biased to higher masses and so the mass distribution more closely follows the IMF. In addition, at the high metallicities in the Milky Way, stellar winds are much stronger and strip away much of the stellar mass before BH formation, resulting the less massive black holes. The mass distribution extends down to $2.5 \unit{M_{\odot}}$, our fiducial maximum neutron star mass. Since the \citet{Fryer+2012} \textit{delayed} remnant mass prescription (applied in this work) does not produce a mass gap between neutron stars and black holes. Indeed we expect $35\%$ and $39\%$ of detected BHBH and BHNS systems to contain a black hole in the lower mass gap. Therefore, LISA could help to confirm or rule out the existence of the lower mass gap.

@@

\subsubsection{Mass Ratio}
The mass ratio distribution for each DCO type is relatively distinct from the others. The majority of NSNSs have a mass ratio close to unity, with 90\% of systems having $q > 0.8$. The reason for the concentration around equal masses is that most NSs are formed either through electron-capture supernovae or from low mass stars. We set the remnant mass for any system formed through ECSN to $1.26 \unit{M_{\odot}}$ (see Sec.~\ref{sec:fiducial_physics}). The remnant mass prescription that we use gives a fixed fallback mass for any star with a CO core mass less than $2.5 \unit{M_\odot}$, such that many NSs are given the identical mass of $1.278 \unit{M_\odot}$ in the \textit{delayed} prescription \citep[see][Eq.~19]{Fryer+2012}. This means that many NSs are formed with equal masses and hence we see a mass ratio distribution peaked around unity.

In contrast, only 9\% of detectable BHBHs are formed with $q > 0.8$ and distribution peaks around $q = 0.4$. The reasoning for these unequal mass ratio systems is as follows: in order to produce a BHBH, most formation channels require at least the first mass transfer to be stable. This stability is strongly dependent on the mass ratio such that equal mass ratios (at the moment of mass transfer) are preferred for creating BHBHs. Yet, since stellar winds are so strong at high metallicity, and even stronger for more massive stars, the primary star will experience significant mass loss and so an initially \textit{unequal} mass ratio is preferred so that the masses are more balanced at the first instance of mass transfer. Since mass transfer occurs after the end of the main sequence for most of our BHBHs, the star will have a well defined core and these core masses, which go on to form BHs, will reflect the initially unequal mass ratios.

@@

An eccentricity of $e = 0.01$ is the lower bound on the measurable eccentricity with LISA proposed by \citet{Nishizawa+2016}. We find that a significant fraction of DCOs (86\% of BHBHs, 46\% of BHNSs and 86\% of NSNSs) exceed this bound. This is in contrast to several previous work that assume all DCOs are circular once they reach the LISA mission \citep[e.g.][]{Lamberts+2018, Sesana+2020}. Moreover, we find that 20\%, 8\% and 12\% of BHBHs, BHNSs and NSNSs respectively have $e > 0.3$. At this eccentricity the majority of the gravitational wave power is emitted in higher harmonics above the orbital frequency and thus the signal would appear at higher frequencies in LISA.

@@

We use Eq.~\ref{eq:chirp_mass_uncertainty} to calculate the chirp mass uncertainty for each DCO in our sample and plot it in Fig.~\ref{fig:m_c_unc}. We find that approximately 10 BHBHs, 10 BHNSs and 5 NSNSs have measurable chirp masses (as indicated by the shaded region). This uncertainty is generally dominated by the uncertainty on the time derivative of the frequency, since most of the binaries are too early in their inspiral for LISA to measure a strong chirp.

>>>>

\subsubsection{Black Hole Mass}
For both the BHBHs and BHNSs, the black hole mass distribution extends to relatively low masses, with $\BHBHmBHBelowEleven{}$ and $\BHNSmBHBelowEleven{}$ respectively below $11 \unit{M_{\odot}}$. Unlike ground-based detectors, LISA is not biased to higher masses and so the mass distribution more closely follows the IMF. In addition, at the high metallicities in the Milky Way, stellar winds are much stronger and strip away much of the stellar mass before BH formation, resulting the less massive black holes. The mass distribution extends down to $2.5 \unit{M_{\odot}}$, our fiducial maximum neutron star mass. Since the \citet{Fryer+2012} \textit{delayed} remnant mass prescription (applied in this work) does not produce a mass gap between neutron stars and black holes. Indeed we expect $\BHBHLowerMassGap{}$ and $\BHNSLowerMassGap{}$ of detected BHBH and BHNS systems to contain a black hole in the lower mass gap. Therefore, LISA could help to confirm or rule out the existence of the lower mass gap.

@@

\subsubsection{Mass Ratio}
The mass ratio distribution for each DCO type is relatively distinct from the others. The majority of NSNSs have a mass ratio close to unity, with $\NSNSqAbovePointEight{}$ of systems having $q > 0.8$. The reason for the concentration around equal masses is that most NSs are formed either through electron-capture supernovae or from low mass stars. We set the remnant mass for any system formed through ECSN to $1.26 \unit{M_{\odot}}$ (see Sec.~\ref{sec:fiducial_physics}). The remnant mass prescription that we use gives a fixed fallback mass for any star with a CO core mass less than $2.5 \unit{M_\odot}$, such that many NSs are given the identical mass of $1.278 \unit{M_\odot}$ in the \textit{delayed} prescription \citep[see][Eq.~19]{Fryer+2012}. This means that many NSs are formed with equal masses and hence we see a mass ratio distribution peaked around unity.

In contrast, only $\BHBHqAbovePointEight{}$ of detectable BHBHs are formed with $q > 0.8$ and distribution peaks around $q = 0.4$. The reasoning for these unequal mass ratio systems is as follows: in order to produce a BHBH, most formation channels require at least the first mass transfer to be stable. This stability is strongly dependent on the mass ratio such that equal mass ratios (at the moment of mass transfer) are preferred for creating BHBHs. Yet, since stellar winds are so strong at high metallicity, and even stronger for more massive stars, the primary star will experience significant mass loss and so an initially \textit{unequal} mass ratio is preferred so that the masses are more balanced at the first instance of mass transfer. Since mass transfer occurs after the end of the main sequence for most of our BHBHs, the star will have a well defined core and these core masses, which go on to form BHs, will reflect the initially unequal mass ratios.

@@

An eccentricity of $e = 0.01$ is the lower bound on the measurable eccentricity with LISA proposed by \citet{Nishizawa+2016}. We find that a significant fraction of DCOs ($\BHBHNotCirc{}$ of BHBHs, $\BHNSNotCirc{}$ of BHNSs and $\NSNSNotCirc{}$ of NSNSs) exceed this bound. This is in contrast to several previous work that assume all DCOs are circular once they reach the LISA mission \citep[e.g.][]{Lamberts+2018, Sesana+2020}. Moreover, we find that $\BHBHHighlyEccentric{}$, $\BHNSHighlyEccentric{}$ and $\NSNSHighlyEccentric{}$ of BHBHs, BHNSs and NSNSs respectively have $e > 0.3$. At this eccentricity the majority of the gravitational wave power is emitted in higher harmonics above the orbital frequency and thus the signal would appear at higher frequencies in LISA.

@@

We use Eq.~\ref{eq:chirp_mass_uncertainty} to calculate the chirp mass uncertainty for each DCO in our sample and plot it in Fig.~\ref{fig:m_c_unc}. We find that approximately $\BHBHChirpMassMeasureable{}$ BHBHs, $\BHNSChirpMassMeasureable{}$ BHNSs and $\NSNSChirpMassMeasureable{}$ NSNSs have measurable chirp masses (as indicated by the shaded region). This uncertainty is generally dominated by the uncertainty on the time derivative of the frequency, since most of the binaries are too early in their inspiral for LISA to measure a strong chirp.
